---
title: "Over Representation Analysis (ORA)"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

ORA can determine if a set of genes shares more or fewer genes with gene sets/pathways than we would expect by chance. Here we look at the orthogroups/genes that have been found to have copy number changes in association with phenotypes (determined by pgls and an FDR cut-off of 0.2).

```{r}
#| echo: false
#| message: false
#| warning: false

setwd("~/OneDrive/Placement/Research/")

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

#BiocManager::install("clusterProfiler")
#BiocManager::install("org.Mm.eg.db")

library(dplyr)
library(clusterProfiler)
library(org.Mm.eg.db)
library(DT)
library(ggplot2)


```

## Data

```{r}
#| echo: false
#| message: false
#| warning: false

## READ IN PGLS RESULTS

maxlong_pos <- read.table("pgls_results/pgls_indv_longevity_pos.txt", header = TRUE, fill = TRUE)
maxlong_neg <- read.table("pgls_results/pgls_indv_longevity_neg.txt", header = TRUE, fill = TRUE)
adultweight_pos <- read.table("pgls_results/pgls_indv_bodysize_pos.txt", header = TRUE, fill = TRUE)
adultweight_neg <- read.table("pgls_results/pgls_indv_bodysize_neg.txt", header = TRUE, fill = TRUE)
neoplasia_pos <- read.table("pgls_results/pgls_indv_neoplasia_pos.txt", header = TRUE, fill = TRUE)
neoplasia_neg<- read.table("pgls_results/pgls_indv_neoplasia_neg.txt", header = TRUE, fill = TRUE)
malignancy_pos <- read.table("pgls_results/pgls_indv_malignancy_pos.txt", header = TRUE, fill = TRUE)
malignancy_neg <- read.table("pgls_results/pgls_indv_malignancy_neg.txt", header = TRUE, fill = TRUE)
ratio_pos <- read.table("pgls_results/pgls_indv_ratio_pos.txt", header = TRUE, fill = TRUE)
ratio_neg <- read.table("pgls_results/pgls_indv_ratio_neg.txt", header = TRUE, fill = TRUE)

#add adjusted p value
adjusted_p_values <- p.adjust(maxlong_pos$p_value, method = "BH")
maxlong_pos$adjusted_p_value <- adjusted_p_values

adjusted_p_values <- p.adjust(maxlong_neg$p_value, method = "BH")
maxlong_neg$adjusted_p_value <- adjusted_p_values

adjusted_p_values <- p.adjust(adultweight_pos$p_value, method = "BH")
adultweight_pos$adjusted_p_value <- adjusted_p_values

adjusted_p_values <- p.adjust(adultweight_neg$p_value, method = "BH")
adultweight_neg$adjusted_p_value <- adjusted_p_values

adjusted_p_values <- p.adjust(neoplasia_pos$p_value, method = "BH")
neoplasia_pos$adjusted_p_value <- adjusted_p_values

adjusted_p_values <- p.adjust(neoplasia_neg$p_value, method = "BH")
neoplasia_neg$adjusted_p_value <- adjusted_p_values

adjusted_p_values <- p.adjust(malignancy_pos$p_value, method = "BH")
malignancy_pos$adjusted_p_value <- adjusted_p_values

adjusted_p_values <- p.adjust(malignancy_neg$p_value, method = "BH")
malignancy_neg$adjusted_p_value <- adjusted_p_values

adjusted_p_values <- p.adjust(ratio_pos$p_value, method = "BH")
ratio_pos$adjusted_p_value <- adjusted_p_values

adjusted_p_values <- p.adjust(ratio_neg$p_value, method = "BH")
ratio_neg$adjusted_p_value <- adjusted_p_values


# add gene names to the results
mouse_ortho <- read.table("mouse_gene.orthogroup_data.tsv", fill = TRUE, header = TRUE)

maxlong_pos <- left_join(mouse_ortho, maxlong_pos, by = c("Orthogroup" = "OG"))
maxlong_neg <- left_join(mouse_ortho, maxlong_neg, by = c("Orthogroup" = "OG"))
adultweight_pos <- left_join(mouse_ortho, adultweight_pos, by = c("Orthogroup" = "OG"))
adultweight_neg <- left_join(mouse_ortho, adultweight_neg, by = c("Orthogroup" = "OG"))
neoplasia_pos <- left_join(mouse_ortho, neoplasia_pos, by = c("Orthogroup" = "OG"))
neoplasia_neg <- left_join(mouse_ortho, neoplasia_neg, by = c("Orthogroup" = "OG"))
malignancy_pos <- left_join(mouse_ortho, malignancy_pos, by = c("Orthogroup" = "OG"))
malignancy_neg <- left_join(mouse_ortho, malignancy_neg, by = c("Orthogroup" = "OG"))
ratio_pos <- left_join(mouse_ortho, ratio_pos, by = c("Orthogroup" = "OG"))
ratio_neg <- left_join(mouse_ortho, ratio_neg, by = c("Orthogroup" = "OG"))

# add NCBI IDs to the results
mouse_ncbi <- read.table("mouse_ncbiIDs_biomart.txt", fill = TRUE, header = TRUE, sep = "\t")

maxlong_pos <- left_join(maxlong_pos, mouse_ncbi, by = c("gene" = "Gene.name"))
maxlong_neg <- left_join(maxlong_neg, mouse_ncbi, by = c("gene" = "Gene.name"))
adultweight_pos <- left_join(adultweight_pos, mouse_ncbi, by = c("gene" = "Gene.name"))
adultweight_neg <- left_join(adultweight_neg, mouse_ncbi, by = c("gene" = "Gene.name"))
neoplasia_pos <- left_join(neoplasia_pos, mouse_ncbi, by = c("gene" = "Gene.name"))
neoplasia_neg <- left_join(neoplasia_neg, mouse_ncbi, by = c("gene" = "Gene.name"))
malignancy_pos <- left_join(malignancy_pos, mouse_ncbi, by = c("gene" = "Gene.name"))
malignancy_neg <- left_join(malignancy_neg, mouse_ncbi, by = c("gene" = "Gene.name"))
ratio_pos <- left_join(ratio_pos, mouse_ncbi, by = c("gene" = "Gene.name"))
ratio_neg <- left_join(ratio_neg, mouse_ncbi, by = c("gene" = "Gene.name"))

# remove orthogroups not found in mouse and orthogroups not in this dataset (pos/neg)
maxlong_pos <- maxlong_pos[complete.cases(maxlong_pos$gene) & maxlong_pos$gene != "" & !is.na(maxlong_pos$coeff) & maxlong_pos$coeff != "", ]
maxlong_neg <- maxlong_neg[complete.cases(maxlong_neg$gene) & maxlong_neg$gene != "" & !is.na(maxlong_neg$coeff) & maxlong_neg$coeff != "", ]
adultweight_pos <- adultweight_pos[complete.cases(adultweight_pos$gene) & adultweight_pos$gene != "" & !is.na(adultweight_pos$coef) & adultweight_pos$coef != "", ]
adultweight_neg <- adultweight_neg[complete.cases(adultweight_neg$gene) & adultweight_neg$gene != "" & !is.na(adultweight_neg$coef) & adultweight_neg$coef != "", ]
neoplasia_pos <- neoplasia_pos[complete.cases(neoplasia_pos$gene) & neoplasia_pos$gene != "" & !is.na(neoplasia_pos$coeff) & neoplasia_pos$coeff != "", ]
neoplasia_neg <- neoplasia_neg[complete.cases(neoplasia_neg$gene) & neoplasia_neg$gene != "" & !is.na(neoplasia_neg$coeff) & neoplasia_neg$coeff != "", ]
malignancy_pos <- malignancy_pos[complete.cases(malignancy_pos$gene) & malignancy_pos$gene != "" & !is.na(malignancy_pos$coeff) & malignancy_pos$coeff != "", ]
malignancy_neg <- malignancy_neg[complete.cases(malignancy_neg$gene) & malignancy_neg$gene != "" & !is.na(malignancy_neg$coeff) & malignancy_neg$coeff != "", ]
ratio_pos <- ratio_pos[complete.cases(ratio_pos$gene) & ratio_pos$gene != "" & !is.na(ratio_pos$coeff) & ratio_pos$coeff != "", ]
ratio_neg <- ratio_neg[complete.cases(ratio_neg$gene) & ratio_neg$gene != "" & !is.na(ratio_neg$coeff) & ratio_neg$coeff != "", ]

#subset for results with p value < 0.05, by using adjusted p.values (Benjamini-Hochberg method)
maxlong_pos.2 <- maxlong_pos[maxlong_pos$p_value < 0.05, ]
maxlong_neg.2 <- maxlong_neg[maxlong_neg$p_value < 0.05, ]
adultweight_pos.2 <- adultweight_pos[adultweight_pos$p_value < 0.05, ]
adultweight_neg.2 <- adultweight_neg[adultweight_neg$p_value < 0.05, ]
neoplasia_pos.2 <- neoplasia_pos[neoplasia_pos$p_value < 0.05, ]
neoplasia_neg.2 <- neoplasia_neg[neoplasia_neg$p_value < 0.05, ]
malignancy_pos.2 <- malignancy_pos[malignancy_pos$p_value < 0.05, ]
malignancy_neg.2 <- malignancy_neg[malignancy_neg$p_value < 0.05, ]
ratio_pos.2 <- ratio_pos[ratio_pos$p_value < 0.05, ]
ratio_neg.2 <- ratio_neg[ratio_neg$p_value < 0.05, ]

```

#### Interpreting ORA results:

The `GeneRatio` column in the enrichment result represents the ratio of the number of genes in your input list that are associated with a particular gene set or term to the total number of genes annotated to that gene set or term.

Interpreting `GeneRatio`:

-   A `GeneRatio` close to 1 indicates that a high proportion of genes from your input list are associated with the gene set or term, suggesting a strong association.

-   A `GeneRatio` significantly smaller than 1 may indicate that the gene set is not well-represented in your input list.

## Maximum longevity (covariate: body size)

### Positive correlations

None of the ORA results for max longevity (positive correlations) have a gene count \> 1.

```{r}
#| echo: false
#| message: false
#| warning: false

# Perform enrichment analysis - positive correlations

#Gene ontology - biological processes
maxlong_pos_gobp <- enrichGO(
  gene          = maxlong_pos.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "BP",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gobp <- as.data.frame(maxlong_pos_gobp)
gobp$group <- "GO: Biological process"
nrow(maxlong_pos_gobp)

#gene ontology - molecular function
maxlong_pos_gomf <- enrichGO(
  gene          = maxlong_pos.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "MF",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gomf <- as.data.frame(maxlong_pos_gomf)
gomf$group <- "GO: Molecular function"
nrow(maxlong_pos_gomf)

#gene ontology - cellular component
maxlong_pos_gocc <- enrichGO(
  gene          = maxlong_pos.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "CC",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gocc <- as.data.frame(maxlong_pos_gocc)
gocc$group <- "GO: Cellular component"
nrow(maxlong_pos_gocc)

#kegg 
maxlong_pos_kegg <- enrichKEGG(
     gene          = maxlong_pos.2$UniProtKB.Gene.Name.ID,
     organism      = 'mmu',  # 'mmu' is the code for Mus musculus (mouse)
     keyType       = "uniprot",
     pvalueCutoff  = 0.05,
     qvalueCutoff  = 0.05,
     pAdjustMethod = "bonferroni"
 )
##keytype: one of "kegg", 'ncbi-geneid', 'ncib-proteinid' and 'uniprot'
kegg <- as.data.frame(maxlong_pos_kegg)
kegg$group <- "KEGG"
nrow(maxlong_pos_kegg)

max_p <- rbind(gomf, gobp, gocc)
#write results for supp data
write.table(max_p, "~/Documents/PhD/Placement/Supp_data/ORA/longevity_pos.txt", quote = FALSE, 
            row.names = FALSE, col.names = TRUE, sep = "\t")

ggplot(max_p, aes(x = GeneRatio, y = Description, color = p.adjust, size = Count)) +
  geom_point() +
  scale_color_continuous() +  # For continuous color scale
  labs(x = "GeneRatio", y = "Description", color = "p.adjust", size = "Count", title = "ORA: Maximum longevity (positive correlations)") +
  theme_bw() +
  facet_grid(group ~ ., scales = "free_y", space = "free_y")
```

### Negative Correlations

```{r}
#| echo: false
#| message: false
#| warning: false

# Perform enrichment analysis - negative correlations

#Gene ontology - biological processes
maxlong_neg_gobp <- enrichGO(
  gene          = maxlong_neg.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "BP",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gobp <- as.data.frame(maxlong_neg_gobp)
gobp$group <- "GO: Biological process"
nrow(maxlong_neg_gobp)

#gene ontology - molecular function
maxlong_neg_gomf <- enrichGO(
  gene          = maxlong_neg.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "MF",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gomf <- as.data.frame(maxlong_neg_gomf)
gomf$group <- "GO: Molecular function"
nrow(maxlong_neg_gomf)

#gene ontology - cellular component
maxlong_neg_gocc <- enrichGO(
  gene          = maxlong_neg.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "CC",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gocc <- as.data.frame(maxlong_neg_gocc)
gocc$group <- "GO: Cellular component"
nrow(maxlong_neg_gocc)

#kegg 
maxlong_neg_kegg <- enrichKEGG(
     gene          = maxlong_neg.2$UniProtKB.Gene.Name.ID,
     organism      = 'mmu',  # 'mmu' is the code for Mus musculus (mouse)
     keyType       = "uniprot",
     pvalueCutoff  = 0.05,
     qvalueCutoff  = 0.05,
     pAdjustMethod = "bonferroni"
 )
##keytype: one of "kegg", 'ncbi-geneid', 'ncib-proteinid' and 'uniprot'
kegg <- as.data.frame(maxlong_neg_kegg)
kegg$group <- "KEGG"
nrow(maxlong_neg_kegg)

max_n <- bind_rows(gomf, gobp, gocc)
#write results for supp data
write.table(max_n, "~/Documents/PhD/Placement/Supp_data/ORA/longevity_neg.txt", quote = FALSE, 
            row.names = FALSE, col.names = TRUE, sep = "\t")

ggplot(max_n, aes(x = GeneRatio, y = Description, color = p.adjust, size = Count)) +
  geom_point() +
  scale_color_continuous() +  # For continuous color scale
  labs(x = "GeneRatio", y = "Description", color = "p.adjust", size = "Count", title = "ORA: Maximum longevity (negative correlations)") +
  theme_bw() +
  facet_grid(group ~ ., scales = "free_y", space = "free_y")
```

```{r}

# combine positive and negative results
max_p$correlation <- "positive"
max_n$correlation <- "negative"

max <- rbind(max_n, max_p)
```

## Adult weight (log transformed w/ covariate: max longevity)

### Positive correlations

```{r}
#| echo: false
#| message: false
#| warning: false

# Perform enrichment analysis - positive correlations

#Gene ontology - biological processes
adultweight_pos_gobp <- enrichGO(
  gene          = adultweight_pos.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "BP",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gobp <- as.data.frame(adultweight_pos_gobp)
gobp$group <- "GO: Biological process"
nrow(adultweight_pos_gobp)

#gene ontology - molecular function
adultweight_pos_gomf <- enrichGO(
  gene          = adultweight_pos.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "MF",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gomf <- as.data.frame(adultweight_pos_gomf)
gomf$group <- "GO: Molecular function"
nrow(adultweight_pos_gomf)

#gene ontology - cellular component
adultweight_pos_gocc <- enrichGO(
  gene          = adultweight_pos.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "CC",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gocc <- as.data.frame(adultweight_pos_gocc)
gocc$group <- "GO: Cellular component"
nrow(adultweight_pos_gocc)

#kegg 
adultweight_pos_kegg <- enrichKEGG(
     gene          = adultweight_pos.2$UniProtKB.Gene.Name.ID,
     organism      = 'mmu',  # 'mmu' is the code for Mus musculus (mouse)
     keyType       = "uniprot",
     pvalueCutoff  = 0.05,
     qvalueCutoff  = 0.05,
     pAdjustMethod = "bonferroni"
 )
##keytype: one of "kegg", 'ncbi-geneid', 'ncib-proteinid' and 'uniprot'
kegg <- as.data.frame(adultweight_pos_kegg)
kegg$group <- "KEGG"
nrow(adultweight_pos_kegg)

size_p <- bind_rows(gomf, gobp, gocc)
#write results for supp data
write.table(size_p, "~/Documents/PhD/Placement/Supp_data/ORA/size_pos.txt", quote = FALSE, 
            row.names = FALSE, col.names = TRUE, sep = "\t")

ggplot(size_p, aes(x = GeneRatio, y = Description, color = p.adjust, size = Count)) +
  geom_point() +
  scale_color_continuous() +  # For continuous color scale
  labs(x = "GeneRatio", y = "Description", color = "p.adjust", size = "Count", title = "ORA: Adult body weight (positive correlations)") +
  theme_bw() +
  facet_grid(group ~ ., scales = "free_y", space = "free_y")
```

### Negative correlations

```{r}
#| echo: false
#| message: false
#| warning: false

# Perform enrichment analysis - negative correlations

#Gene ontology - biological processes
adultweight_neg_gobp <- enrichGO(
  gene          = adultweight_neg.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "BP",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gobp <- as.data.frame(adultweight_neg_gobp)
gobp$group <- "GO: Biological process"
nrow(maxlong_neg_gobp)

#gene ontology - molecular function
maxlong_neg_gomf <- enrichGO(
  gene          = maxlong_neg.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "MF",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gomf <- as.data.frame(maxlong_neg_gomf)
gomf$group <- "GO: Molecular function"
nrow(maxlong_neg_gomf)

#gene ontology - cellular component
maxlong_neg_gocc <- enrichGO(
  gene          = maxlong_neg.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "CC",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gocc <- as.data.frame(maxlong_neg_gocc)
gocc$group <- "GO: Cellular component"
nrow(maxlong_neg_gocc)

#kegg 
maxlong_neg_kegg <- enrichKEGG(
     gene          = maxlong_neg.2$UniProtKB.Gene.Name.ID,
     organism      = 'mmu',  # 'mmu' is the code for Mus musculus (mouse)
     keyType       = "uniprot",
     pvalueCutoff  = 0.05,
     qvalueCutoff  = 0.05,
     pAdjustMethod = "bonferroni"
 )
##keytype: one of "kegg", 'ncbi-geneid', 'ncib-proteinid' and 'uniprot'
kegg <- as.data.frame(maxlong_neg_kegg)
#kegg$group <- "KEGG"
nrow(maxlong_neg_kegg)

size_n <- bind_rows(gomf, gobp, gocc)
#write results for supp data
write.table(size_n, "~/Documents/PhD/Placement/Supp_data/ORA/size_neg.txt", quote = FALSE, 
            row.names = FALSE, col.names = TRUE, sep = "\t")

ggplot(size_n, aes(x = GeneRatio, y = Description, color = p.adjust, size = Count)) +
  geom_point() +
  scale_color_continuous() +  # For continuous color scale
  labs(x = "GeneRatio", y = "Description", color = "p.adjust", size = "Count", title = "ORA: Adult body weight (negative correlations)") +
  theme_bw() +
  facet_grid(group ~ ., scales = "free_y", space = "free_y")
```

```{r}

# combine positive and negative results
size_p$correlation <- "positive"
size_n$correlation <- "negative"

size <- rbind(size_n, size_p)
```

## Neoplasia prevalence

### Positive correlations

```{r}
#| echo: false
#| message: false
#| warning: false

# Perform enrichment analysis - positive correlations

#Gene ontology - biological processes
neoplasia_pos_gobp <- enrichGO(
  gene          = neoplasia_pos.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "BP",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gobp <- as.data.frame(neoplasia_pos_gobp)
gobp$group <- "GO: Biological process"
nrow(neoplasia_pos_gobp)

#gene ontology - molecular function
neoplasia_pos_gomf <- enrichGO(
  gene          = neoplasia_pos.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "MF",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gomf <- as.data.frame(neoplasia_pos_gomf)
gomf$group <- "GO: Molecular function"
nrow(neoplasia_pos_gomf)

#gene ontology - cellular component
neoplasia_pos_gocc <- enrichGO(
  gene          = neoplasia_pos.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "CC",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gocc <- as.data.frame(neoplasia_pos_gocc)
gocc$group <- "GO: Cellular component"
nrow(neoplasia_pos_gocc)

#kegg 
neoplasia_pos_kegg <- enrichKEGG(
     gene          = neoplasia_pos.2$UniProtKB.Gene.Name.ID,
     organism      = 'mmu',  # 'mmu' is the code for Mus musculus (mouse)
     keyType       = "uniprot",
     pvalueCutoff  = 0.05,
     qvalueCutoff  = 0.05,
     pAdjustMethod = "bonferroni"
 )
##keytype: one of "kegg", 'ncbi-geneid', 'ncib-proteinid' and 'uniprot'
kegg <- as.data.frame(neoplasia_pos_kegg)
kegg$group <- "KEGG"
nrow(neoplasia_pos_kegg)

neo_p <- bind_rows(gomf, gobp, gocc)
#write results for supp data
write.table(neo_p, "~/Documents/PhD/Placement/Supp_data/ORA/neo_pos.txt", quote = FALSE, 
            row.names = FALSE, col.names = TRUE, sep = "\t")

ggplot(neo_p, aes(x = GeneRatio, y = Description, color = p.adjust, size = Count)) +
  geom_point() +
  scale_color_continuous() +  # For continuous color scale
  labs(x = "GeneRatio", y = "Description", color = "p.adjust", size = "Count", title = "ORA: Neoplasia prevalence (positive correlations)") +
  theme_bw() +
  facet_grid(group ~ ., scales = "free_y", space = "free_y")
```

### Negative correlations

```{r}
#| echo: false
#| message: false
#| warning: false

# Perform enrichment analysis - negative correlations

#Gene ontology - biological processes
neoplasia_neg_gobp <- enrichGO(
  gene          = neoplasia_neg.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "BP",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gobp <- as.data.frame(neoplasia_neg_gobp)
gobp$group <- "GO: Biological process"
nrow(neoplasia_neg_gobp)

#gene ontology - molecular function
neoplasia_neg_gomf <- enrichGO(
  gene          = neoplasia_neg.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "MF",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gomf <- as.data.frame(neoplasia_neg_gomf)
gomf$group <- "GO: Molecular function"
nrow(neoplasia_neg_gomf)

#gene ontology - cellular component
neoplasia_neg_gocc <- enrichGO(
  gene          = neoplasia_neg.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "CC",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gocc <- as.data.frame(neoplasia_neg_gocc)
gocc$group <- "GO: Cellular component"
nrow(neoplasia_neg_gocc)

#kegg 
neoplasia_neg_kegg <- enrichKEGG(
     gene          = neoplasia_neg.2$UniProtKB.Gene.Name.ID,
     organism      = 'mmu',  # 'mmu' is the code for Mus musculus (mouse)
     keyType       = "uniprot",
     pvalueCutoff  = 0.05,
     qvalueCutoff  = 0.05,
     pAdjustMethod = "bonferroni"
 )
##keytype: one of "kegg", 'ncbi-geneid', 'ncib-proteinid' and 'uniprot'
kegg <- as.data.frame(neoplasia_neg_kegg)
kegg$group <- "KEGG"
nrow(neoplasia_neg_kegg)

neo_n <- bind_rows(gomf, gobp, gocc)
#write results for supp data
write.table(neo_n, "~/Documents/PhD/Placement/Supp_data/ORA/neo_neg.txt", quote = FALSE, 
            row.names = FALSE, col.names = TRUE, sep = "\t")

## make the Generatio column an integer
# Split the fractions into numerator and denominator
fraction_parts <- strsplit(neo_n$GeneRatio, "/")
# Convert numerator and denominator to numeric
numerator <- as.numeric(sapply(fraction_parts, "[[", 1))
denominator <- as.numeric(sapply(fraction_parts, "[[", 2))
# Convert fractions to decimals
neo_n$GeneRatio <- numerator / denominator

p <- ggplot(neo_n, aes(x = GeneRatio, y = Description, color = p.adjust, size = Count)) +
  geom_point() +
  scale_color_continuous() +  # For continuous color scale
  labs(x = "GeneRatio", y = "Ontology term", color = "adjus", size = "Count", title = "ORA: Neoplasia prevalence (negative correlations)") +
  theme_bw() +
  facet_grid(group ~ ., scales = "free_y", space = "free_y")
```

```{r}

# combine positive and negative results
neo_p$correlation <- "positive"
neo_n$correlation <- "negative"

neo <- rbind(neo_n, neo_p)
```

## Malignancy prevalence

### Positive correlations

```{r}
#| echo: false
#| message: false
#| warning: false

# Perform enrichment analysis - positive correlations

#Gene ontology - biological processes
malignancy_pos_gobp <- enrichGO(
  gene          = malignancy_pos.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "BP",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gobp <- as.data.frame(malignancy_pos_gobp)
gobp$group <- "GO: Biological process"
nrow(malignancy_pos_gobp)

#gene ontology - molecular function
malignancy_pos_gomf <- enrichGO(
  gene          = malignancy_pos.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "MF",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gomf <- as.data.frame(malignancy_pos_gomf)
gomf$group <- "GO: Molecular function"
nrow(malignancy_pos_gomf)

#gene ontology - cellular component
malignancy_pos_gocc <- enrichGO(
  gene          = malignancy_pos.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "CC",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gocc <- as.data.frame(malignancy_pos_gocc)
gocc$group <- "GO: Cellular component"
nrow(malignancy_pos_gocc)

#kegg 
malignancy_pos_kegg <- enrichKEGG(
     gene          = malignancy_pos.2$UniProtKB.Gene.Name.ID,
     organism      = 'mmu',  # 'mmu' is the code for Mus musculus (mouse)
     keyType       = "uniprot",
     pvalueCutoff  = 0.05,
     qvalueCutoff  = 0.05,
     pAdjustMethod = "bonferroni"
 )
##keytype: one of "kegg", 'ncbi-geneid', 'ncib-proteinid' and 'uniprot'
kegg <- as.data.frame(malignancy_pos_kegg)
kegg$group <- "KEGG"
nrow(malignancy_pos_kegg)

mal_p <- bind_rows(gomf, gobp, gocc)
#write results for supp data
write.table(mal_p, "~/Documents/PhD/Placement/Supp_data/ORA/mal_pos.txt", quote = FALSE, 
            row.names = FALSE, col.names = TRUE, sep = "\t")

ggplot(mal_p, aes(x = GeneRatio, y = Description, color = p.adjust, size = Count)) +
  geom_point() +
  scale_color_continuous() +  # For continuous color scale
  labs(x = "GeneRatio", y = "Description", color = "p.adjust", size = "Count", title = "ORA: Malignancy prevalence (positive correlations)") +
  theme_bw() +
  facet_grid(group ~ ., scales = "free_y", space = "free_y")
```

### Negative correlations

```{r}
#| echo: false
#| message: false
#| warning: false

# Perform enrichment analysis - negative correlations

#Gene ontology - biological processes
malignancy_neg_gobp <- enrichGO(
  gene          = malignancy_neg.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "BP",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gobp <- as.data.frame(malignancy_neg_gobp)
gobp$group <- "GO: Biological process"
nrow(malignancy_neg_gobp)

#gene ontology - molecular function
malignancy_neg_gomf <- enrichGO(
  gene          = malignancy_neg.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "MF",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gomf <- as.data.frame(malignancy_neg_gomf)
gomf$group <- "GO: Molecular function"
nrow(malignancy_neg_gomf)

#gene ontology - cellular component
malignancy_neg_gocc <- enrichGO(
  gene          = malignancy_neg.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "CC",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gocc <- as.data.frame(malignancy_neg_gocc)
gocc$group <- "GO: Cellular component"
nrow(malignancy_neg_gocc)

#kegg 
malignancy_neg_kegg <- enrichKEGG(
     gene          = malignancy_neg.2$UniProtKB.Gene.Name.ID,
     organism      = 'mmu',  # 'mmu' is the code for Mus musculus (mouse)
     keyType       = "uniprot",
     pvalueCutoff  = 0.05,
     qvalueCutoff  = 0.05,
     pAdjustMethod = "bonferroni"
 )
##keytype: one of "kegg", 'ncbi-geneid', 'ncib-proteinid' and 'uniprot'
kegg <- as.data.frame(malignancy_neg_kegg)
kegg$group <- "KEGG"
nrow(malignancy_neg_kegg)

mal_n <- bind_rows(gomf, gobp, gocc)
#write results for supp data
write.table(mal_n, "~/Documents/PhD/Placement/Supp_data/ORA/mal_neg.txt", quote = FALSE, 
            row.names = FALSE, col.names = TRUE, sep = "\t")

ggplot(mal_n, aes(x = GeneRatio, y = Description, color = p.adjust, size = Count)) +
  geom_point() +
  scale_color_continuous() +  # For continuous color scale
  labs(x = "GeneRatio", y = "Description", color = "p.adjust", size = "Count", title = "ORA: Malignancy prevalence (negative correlations)") +
  theme_bw() +
  facet_grid(group ~ ., scales = "free_y", space = "free_y")
```

```{r}

# combine positive and negative results
mal_p$correlation <- "positive"
mal_n$correlation <- "negative"

mal <- rbind(mal_n, mal_p)
```

## Neoplasia:Malignancy ratio

### Positive correlations

```{r}
#| echo: false
#| message: false
#| warning: false

# Perform enrichment analysis - positive correlations

#Gene ontology - biological processes
ratio_pos_gobp <- enrichGO(
  gene          = ratio_pos.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "BP",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gobp <- as.data.frame(ratio_pos_gobp)
gobp$group <- "GO: Biological process"
nrow(ratio_pos_gobp)

#gene ontology - molecular function
ratio_pos_gomf <- enrichGO(
  gene          = ratio_pos.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "MF",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gomf <- as.data.frame(ratio_pos_gomf)
gomf$group <- "GO: Molecular function"
nrow(ratio_pos_gomf)

#gene ontology - cellular component
ratio_pos_gocc <- enrichGO(
  gene          = ratio_pos.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "CC",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gocc <- as.data.frame(ratio_pos_gocc)
gocc$group <- "GO: Cellular component"
nrow(ratio_pos_gocc)

#kegg 
ratio_pos_kegg <- enrichKEGG(
     gene          = ratio_pos.2$UniProtKB.Gene.Name.ID,
     organism      = 'mmu',  # 'mmu' is the code for Mus musculus (mouse)
     keyType       = "uniprot",
     pvalueCutoff  = 0.05,
     qvalueCutoff  = 0.05,
     pAdjustMethod = "bonferroni"
 )
##keytype: one of "kegg", 'ncbi-geneid', 'ncib-proteinid' and 'uniprot'
kegg <- as.data.frame(ratio_pos_kegg)
kegg$group <- "KEGG"
nrow(ratio_pos_kegg)

rat_p <- bind_rows(gomf, gobp, gocc)
#write results for supp data
write.table(rat_p, "~/Documents/PhD/Placement/Supp_data/ORA/ratio_pos.txt", quote = FALSE, 
            row.names = FALSE, col.names = TRUE, sep = "\t")

ggplot(rat_p, aes(x = GeneRatio, y = Description, color = p.adjust, size = Count)) +
  geom_point() +
  scale_color_continuous() +  # For continuous color scale
  labs(x = "GeneRatio", y = "Description", color = "p.adjust", size = "Count", title = "ORA: Malignancy/neoplasia ratio (positive correlations)") +
  theme_bw() +
  facet_grid(group ~ ., scales = "free_y", space = "free_y",
             labeller = label_wrap_gen(width=10))
```

### Negative correlations

```{r}
#| echo: false
#| message: false
#| warning: false

# Perform enrichment analysis - negative correlations

#Gene ontology - biological processes
ratio_neg_gobp <- enrichGO(
  gene          = ratio_neg.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "BP",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gobp <- as.data.frame(ratio_neg_gobp)
gobp$group <- "GO: Biological process"
nrow(ratio_neg_gobp)

#gene ontology - molecular function
ratio_neg_gomf <- enrichGO(
  gene          = ratio_neg.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "MF",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gomf <- as.data.frame(ratio_neg_gomf)
gomf$group <- "GO: Molecular function"
nrow(ratio_neg_gomf)

#gene ontology - cellular component
ratio_neg_gocc <- enrichGO(
  gene          = ratio_neg.2$gene,
  OrgDb         = org.Mm.eg.db,
  keyType       = "SYMBOL",
  ont           = "CC",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  pAdjustMethod = "bonferroni"
)
gocc <- as.data.frame(ratio_neg_gocc)
gocc$group <- "GO: Cellular component"
nrow(ratio_neg_gocc)

#kegg 
ratio_neg_kegg <- enrichKEGG(
     gene          = ratio_neg.2$UniProtKB.Gene.Name.ID,
     organism      = 'mmu',  # 'mmu' is the code for Mus musculus (mouse)
     keyType       = "uniprot",
     pvalueCutoff  = 0.05,
     qvalueCutoff  = 0.05,
     pAdjustMethod = "bonferroni"
 )
##keytype: one of "kegg", 'ncbi-geneid', 'ncib-proteinid' and 'uniprot'
kegg <- as.data.frame(ratio_neg_kegg)
kegg$group <- "KEGG"
nrow(ratio_neg_kegg)

rat_n <- bind_rows(gomf, gobp, gocc)
#write results for supp data
write.table(rat_n, "~/Documents/PhD/Placement/Supp_data/ORA/rat_neg.txt", quote = FALSE, 
            row.names = FALSE, col.names = TRUE, sep = "\t")

ggplot(rat_n, aes(x = GeneRatio, y = Description, color = p.adjust, size = Count)) +
  geom_point() +
  scale_color_continuous() +  # For continuous color scale
  labs(x = "GeneRatio", y = "Description", color = "p.adjust", size = "Count", title = "ORA: Malignancy/neoplasia ratio (negative correlations)") +
  theme_bw() +
  facet_grid(group ~ ., scales = "free_y", space = "free_y",
             labeller = label_wrap_gen(width=10)) +
  scale_x_continuous()
```

```{r}

# combine positive and negative results
rat_p$correlation <- "positive"
rat_n$correlation <- "negative"

rat <- rbind(rat_n, rat_p)

write.table(rat, "ORAresults_ratio.txt", quote=FALSE, row.names = FALSE, col.names = TRUE, sep = "\t")
```
